---
title: "covid 19"
author: "Michael Maida"
date: "Fall 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + sex,data=kRdata)
```

Based on this graph, it is immediately obvious that there are significantly more females than there are males. If you stacked both of the blue bars and compared them to both of the pink bars, the blue bar, which represent the women who participated in this trial, is far taller than the pink. 

```{r}
barchartGC(~result + sex,data=kRdata, type="percent")
```

This second graph indicates that there is a larger percent of females living and a significantly larger percent of men who are dead. It seems as though 57% of the living population is female while 43% are male, and that 63% of the deceased population is male while 37% are female. This inverse proportionality makes  sense in light of the fact that there are far more females than males in this test population.


  
## Numerical results


## Inferential results 




## Conclusions

